What does the crowd believe? A hierarchical approach to estimating subjective beliefs from empirical data

### 2016-01-29 text/01_CogSci_abstract/paper.tex
edits

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{subcaption}
\usepackage{todonotes}
\usepackage{amsmath}

@@

\begin{abstract}
  Subjects' beliefs about everyday events are of theoretical interest on their own and also an
  important ingredient in especially Bayesian models of such diverse phenomena as logical
  reasoning, future predictions or language use and interpretation. Here, we scrutinize one
  recently popular method for measuring subjective beliefs experimentally. We present a
  hierarchical Bayesian model for inferring likely ``population-level beliefs'' as the central
  tendency of subjects' individual-level beliefs. Individual-level beliefs define likelihood
  functions for three types of task measures. Our results suggest that the previous practice of
  using averaged normalized slider ratings for binned quantities is a practical and fairly good
  approximator of latently inferred population-level beliefs.

\textbf{Keywords:} subjective beliefs, hierarchical modeling, Bayesian data analysis, Bayesian
cognitive models 
\end{abstract}

@@

We cannot look into a person's head. But in making sense of observed behavior we readily
ascribe beliefs and desires to fellow agents. This happens intuitively, in folk psychology, but
also in science. Ascriptions of latent mental states play an important role as well in many
explanations of especially higher-order cognition, like decision making, planning, reasoning,
or language use. Naturally, then, it becomes important how to validate any putative ascription
of mental states for explanatory purposes.

A family of models where this is particularly pressing are Bayesian models of cognition which
seek to explain task behavior in a variety of domains as partially informed by what subjects
believe about mundane events. Take interpretation of language. Empirical data on whether a
statement like ``That watch cost $n$ dollars'' is understood to convey speaker affect, rather
than literal meaning, can be explained well by a Bayesian model of utterance interpretation
\citep{KaoWu2014:Nonliteral-Unde} in which a crucial role is played, as is quite intuitive, by
an empirical measure of subjects' expectations about the likely or normal price of a
watch. Other examples of domains in which empirically successful models have included some
measure of subjects' belief include making future predictions
\citep{GriffithsTenenbaum2006:Optimal-Predict}, the strength of pragmatic enrichments
\citep{DegenTessler2015:Wonky-worlds:-L}, the interpretation of vague quantifiers
\citep{SchollerFranke2015:Semantic-values}, \textcolor{red}{ \dots }

@@

Many methods of assessing subjective beliefs exist. One approach is to take actual frequencies
as an approximation \citep{GriffithsTenenbaum2006:Optimal-Predict}. This, however, does not
work for beliefs about one-shot or imaginary events. Moreover, even where available, real-world
frequencies may deviate systematically from subjects' beliefs, and this can be problematic for
the predictions of a model that relies on subjective beliefs
\citep{MarcusDavis2013:How-Robust-Are-}.

Another approach is to try to experimentally measure relevant beliefs. Many techniques for this
exist, especially in the economics literature
\citep{MorganHenrion1990:Uncertainty:-A-,Manski2004:Measuring-Expec,SchlagTremewan2014:A-penny-for-you,AndersenFountain2014:Estimating-Subj}. The
perhaps most prominent method uses so-called \emph{scoring rules}
\citep{Savage1971:Elicitation-of-,SchlagTremewan2014:A-penny-for-you}. Roughly speaking,
scoring rules are particular schemes of rewarding participants' answers in such a way as to
make sure that responses are what we want them to be: the actual subjective probability of an
event, the mean of a distribution, a credible interval etc. Scoring rules are highly faithful
but require participants to become sufficiently familiar with the (usually probabilistic)
payoff scheme. Scoring rules also must make assumptions about how much participants value
certain probabilistic gambles and may therefore require elaborate means of data analysis
\citep{AndersenFountain2014:Estimating-Subj}. Similar practical concerns apply to other methods
as well, such as the use of \emph{iterated learning} for prior estimation
\citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.

@@

@@

\bibliography{MyRefGlobal}

>>>>

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{subcaption}
\usepackage{todonotes}
\usepackage{amsmath}
\usepackage{color}

\definecolor{Green}{RGB}{10,200,100}
\newcommand{\ndg}[1]{\textcolor{Green}{[ndg: #1]}}

@@

\begin{abstract}
  Subjects' beliefs about everyday events are of theoretical interest on their own and also an
  important ingredient for model building---especially in Bayesian models of phenomena such as logical
  reasoning, future predictions, and language use. Here, we explore several recently used 
  methods for measuring subjective beliefs about unidimensional properties. \ndg{say what kind of subjective beliefs /distributions we look at more clearly.} 
  We a hierarchical Bayesian data analysis model for inferring likely ``population-level beliefs'' as the central
  tendency of subjects' individual-level beliefs. 
  Three different dependent measures---\ndg{name them?}---are related to individual-level beliefs through likelihood ``linking functions.''
%  functions for three types of task measures. 
  Our results suggest that the practice of
  using averaged normalized slider ratings for binned quantities is a practical and fairly good
  approximator of inferred population-level beliefs.
  Other dependent measures yield less reliable results.
\textbf{Keywords:} subjective beliefs, hierarchical modeling, Bayesian data analysis, Bayesian
cognitive models 
\end{abstract}

@@

We cannot look into a person's head. But in making sense of observed behavior we readily
ascribe beliefs and desires to fellow agents. This happens intuitively, in folk psychology, but
also in science. Scientific ascription of latent mental states plays an important role in many
explanations of higher-order cognition: decision making, planning, reasoning,
and language use. Naturally, then, it becomes important to validate any method for eliciting mental states, 
to be used for explanatory purposes.

A family of models where this is particularly pressing are Bayesian models of cognition which
seek to explain task behavior in a variety of domains as partially informed by what subjects
believe about mundane events---their \emph{prior beliefs}. Take interpretation of language. Empirical data on whether a
statement like ``That watch cost $n$ dollars'' is understood to convey speaker affect, rather
than literal meaning, can be explained well by a Bayesian model of utterance interpretation
\citep{KaoWu2014:Nonliteral-Unde}. In this model a crucial role is played by
an empirical measure of subjects' expectations about the likely or normal price of a
watch. 
Other examples of domains in which empirically successful models have included some
measure of subjects' belief include making future predictions
\citep{GriffithsTenenbaum2006:Optimal-Predict}, the strength of pragmatic enrichments
\citep{DegenTessler2015:Wonky-worlds:-L}, the interpretation of vague quantifiers
\citep{SchollerFranke2015:Semantic-values}, \textcolor{red}{ \dots }

@@

Many methods of assessing subjective beliefs exist. One approach is to take actual frequencies
as an approximation \citep{GriffithsTenenbaum2006:Optimal-Predict}. This will not
work when sufficient real-world data is unavailable---for instance, for beliefs about imaginary events. 
Even where available, real-world
frequencies may deviate systematically from subjects' beliefs, and this can be problematic for
the predictions of a model that relies on subjective beliefs.
%\citep{MarcusDavis2013:How-Robust-Are-}.

Another approach is to experimentally elicit relevant beliefs. Many techniques for this
exist, especially in the economics literature
\citep{MorganHenrion1990:Uncertainty:-A-,Manski2004:Measuring-Expec,SchlagTremewan2014:A-penny-for-you,AndersenFountain2014:Estimating-Subj}. The
perhaps most prominent method uses so-called \emph{scoring rules}
\citep{Savage1971:Elicitation-of-,SchlagTremewan2014:A-penny-for-you}. Roughly speaking,
scoring rules are particular schemes of rewarding participants' answers in such a way as to
make sure that responses are what we want them to be: the actual subjective probability of an
event, the mean of a distribution, a credible interval etc. Scoring rules are highly faithful
but require participants to become sufficiently familiar with the (usually probabilistic)
payoff scheme. Scoring rules also must make assumptions about how much participants value
certain probabilistic gambles and may therefore require elaborate means of data analysis
\citep{AndersenFountain2014:Estimating-Subj}. Similar practical concerns apply to other methods
as well, such as the use of \emph{iterated learning} for prior estimation
\citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.

@@

\ndg{should we also talk about the other dependent measures? i think part of what's interesting is comparing the different dependent measures to each other?}

\ndg{say more clearly that we focus on unidimensional distributions and why.}

@@

\bibliography{references}
